\texttt{thornado}-transport: IMEX schemes for two-moment neutrino transport respecting Fermi-Dirac statistics

### 2018-10-17 Documents/M1/Astronum_2018/manuscript.tex
Added acknowledgment. Cosmetic changes for the sine wave tests plots.
Updated the Astronum_2018 document.

\begin{abstract}
We develop implicit-explicit (IMEX) schemes for neutrino transport though a background material in \texttt{thornado} code with two-moment method that evolves the angular moments of a neutrino phase space distribution function.
Considering the upper and lower bounds that introduced by Pauli's exclusion principle on the moments, a algebraic moment closure based on Fermi-Dirac statistics and a convex-invariant time integrator are demanded.
Severn algebraic closures are compared with these bounds and Cernohorsky \& Bludman closure\cite{cernohorskyBludman_1994} which preserves the bounds is chosen for our IMEX schemes.
For the convex-invariant time integrator, two IMEX scheme named PD-ARSs have been proposed following the work in \cite{chu_etal_2018}.
PD-ARS denotes convex-invariant IMEX Runge-Kutta scheme that has first-order accuracy, high-order accuracy in the streaming limit, and works well in the diffusion limit.
Our two PD-ARSs have second-order and third-order explicit strong stability-preserving Runge-Kutta methods\cite{gottlieb_etal_2001} as their explicit part, respectively, and therefore are second-order and third-order accurate in the streaming limit.
The accuracy and convex-invariance of PD-ARSs are demonstrated in the numerical tests with a first-order discontinuous Galerkin(DG) method for spatial discretization and a simple Lax-Friedrichs flux.

@@

\input{sections/intro}
\input{sections/TwoMoments}
\input{sections/SpacialDiscretization}
\input{sections/TimeIntegration}
\input{sections/Numerical}
\input{sections/Conclusion}

\section*{References}
\bibliography{references/references}

>>>>

\begin{abstract}
We develop implicit-explicit (IMEX) schemes for neutrino transport through a background material in the \texttt{thornado} code in the context of a two-moment mode that evolves the angular moments of a neutrino phase-space distribution function.
Considering the upper and lower bounds that are introduced by Pauli's exclusion principle on the moments, an algebraic moment closure based on Fermi-Dirac statistics and a convex-invariant time integrator are demanded.
Several algebraic closures are compared with these bounds in mind and the Cernohorsky and Bludman closure\cite{cernohorskyBludman_1994}, which preserves the bounds, is chosen for our IMEX schemes.
For the convex-invariant time integrator, two IMEX scheme named PD-ARS have been proposed following the work in \cite{chu_etal_2018}.
PD-ARS denotes a convex-invariant IMEX Runge-Kutta scheme that has first-order accuracy in general, high-order accuracy in the streaming limit, and works well in the diffusion limit.
Our two PD-ARS schemes are second- and third-order, explicit, strong-stability-preserving Runge-Kutta methods\cite{gottlieb_etal_2001} as their explicit part, respectively, and therefore are second- and third-order accurate in the streaming limit, respectively.
The accuracy and convex-invariance of our PD-ARS schemes are demonstrated in the numerical tests with a first-order discontinuous Galerkin(DG) method for spatial discretization and a simple Lax-Friedrichs flux.

@@

\input{sections/intro}
\input{sections/TwoMoments}
\input{sections/SpacialDiscretization}
\input{sections/TimeIntegration}
\input{sections/Numerical}
\input{sections/Conclusion}
\input{sections/Acknowledge}

%\section*{References}
\bibliography{references/references}

### 2018-10-17 Documents/M1/Astronum_2018/sections/Acknowledge.tex
Updated the Astronum_2018 document.

\rc{This research is sponsored, in part, by the Laboratory Directed Research and Development Program of Oak Ridge National Laboratory (ORNL), managed by UT-Battelle, LLC for the U. S. Department of Energy under Contract No. De-AC05-00OR22725.  
This research was supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy Office of Science and the National Nuclear Security Administration.  
This material is based, in part, upon work supported by the U.S. Department of Energy, Office of Science, Office of Advanced Scientific Computing Research.  
Ran Chu was supported in part by NSF under Grant No. 1535130.}

>>>>

This research is sponsored, in part, by the Laboratory Directed Research and Development Program of Oak Ridge National Laboratory (ORNL), managed by UT-Battelle, LLC for the U. S. Department of Energy under Contract No. De-AC05-00OR22725.  
This research was supported by the Exascale Computing Project (17-SC-20-SC), a collaborative effort of the U.S. Department of Energy Office of Science and the National Nuclear Security Administration.  
This material is based, in part, upon work supported by the U.S. Department of Energy, Office of Science, Office of Advanced Scientific Computing Research.
Eirik Endeve and Anthony Mezzacappa acknowledge support from the National Science Foundation under group NSF-GP-... and from the UT-ORNL Joint Institute for Computational Sciences.

### 2018-10-17 Documents/M1/Astronum_2018/sections/Conclusion.tex
Updated the Astronum_2018 document.

We have developed IMEX schemes for the two-moment neutrino transport in \texttt{thornado} code respecting Fermi-Dirac statistics.
The schemes employ algebraic closure based on Fermi-Dirac statistics, a first-order discontinuous Galerkin method, the simple Lax-Friedrichs flux with the convex-invariant time integrator to maintain point-wise realizability of the moments.
Since the realizability-preserving property is obtained from convexity of the realizable domain, it's possible to construct a scheme with high order DG method.
See \cite{chu_etal_2018} for some examples with third-order DG method employed.

In the applications motivating this work, the neutrino distribution function can vary from 0 to 1.
Hence, we have considered the algebraic closure respecting Fermi-Dirac statistics for both low and high occupied condition: different values given by different algebraic closures are compared with the boundary values given by the Fermi-Dirac statistics at different occupation.
Among the seven algebraic closures we considered, Kershaw\cite{kershaw_1976}, Wilson\cite{wilson_1975,leblancWilson_1970}, Levermore\cite{levermore_1984}, Minerbo \cite{minerbo_1978}, Janka 1\cite{janka_1991} Janka 2\cite{janka_1992} and Cernohorsky \& Bludman \cite{cernohorskyBludman_1994}, only Cernohorsky \& Bludman closure respecting Fermi-Dirac statistics for all the occupancy.
So we employ Cernohorsky \& Bludman closure in all the numerical tests in Section~\ref{se:NumericalTests}.

Two PD-ARSs are proposed for the desired IMEX scheme.
The one with SSPRK2 is second-order and strong stability-preserving in the streaming limit while other with SSPRK3 is third-order.
Their properties, both accuracy and convex-invariant, are demonstrated with numerical tests.

In this work, we adopted Cartesian coordinates, linear collision term, non-relativistic, and fixed material background.
More realistic problems of scientific interest, such as with the scattering with energy changes and relativistic effects, are left for future research.

>>>>

We have developed IMEX schemes for two-moment neutrino transport in \texttt{thornado} that respect Fermi-Dirac statistics.
The schemes employ algebraic closure based on Fermi-Dirac statistics, a first-order discontinuous Galerkin method, the simple Lax-Friedrichs flux, and convex-invariant time integration to maintain point-wise realizability of the moments.
Since the realizability-preserving property is obtained from the convexity of the realizable domain, it's possible to construct a scheme with a high-order DG method.
See \cite{chu_etal_2018} for some examples with third-order DG methods.

In the application motivating this work, the neutrino distribution function can vary from 0 to 1.
Hence, we have considered algebraic closure respecting Fermi-Dirac statistics for both low and high occupancy.
Among the seven algebraic closures we considered -- Kershaw\cite{kershaw_1976}, Wilson\cite{wilson_1975,leblancWilson_1970}, Levermore\cite{levermore_1984}, Minerbo \cite{minerbo_1978}, Janka 1\cite{janka_1991}, Janka 2\cite{janka_1992}, and Cernohorsky \& Bludman\cite{cernohorskyBludman_1994} -- only the Cernohorsky \& Bludman closure respected Fermi-Dirac statistics for all occupancies.
As a result, we employed the Cernohorsky \& Bludman closure in all of the numerical tests in Section~\ref{se:NumericalTests}.

Two PD-ARS schemes are proposed for the desired IMEX scheme.
The one with SSPRK2 is second-order accurate and strong-stability preserving in the streaming limit while the other with SSPRK3 is third-order accurate.
Their properties, both accuracy and convex-invariance, were demonstrated with numerical tests.

In this work, we adopted Cartesian coordinates, linear collision term, and a fixed non-relativistic material background.
More realistic problems of scientific interest, such as with the scattering with energy exchange and relativistic effects, are left for future research.

### 2018-10-17 Documents/M1/Astronum_2018/sections/Numerical.tex
Added acknowledgment. Cosmetic changes for the sine wave tests plots.
Updated the Astronum_2018 document.

In this section we present the numerical results obtained with the PD-ARS schemes provided in this paper.
The tests in Section \ref{se: Accuracy Tests} are designed to compare the accuracy of the schemes in various regimes.
The test in Section \ref{se: Neutrino Stationary State Test} demonstrates the realizability-preserving of PD-ARSs scheme.
All the tests in this subsection were applied with third-order accurate spatial discretization (polynomials of degree $k=2$) and time step $\dt = 0.1 \times \dx $.
\subsection{Accuracy Tests}
\label{se: Accuracy Tests}
To compare the accuracy of the IMEX scheme, we applied our PDARSs and Pareschi \& Russo \cite{pareschiRusso_2005} (SSP2332) scheme to some smooth problems in streaming, absorption, ans scattering dominated regimes in one spatial dimension.
All the tests in this subsection were applied with the maximum entropy closure in the low occupancy limit.
In the streaming test, the second-order and third-order accurate explicit strong stability-preserving Runge-Kutta methods\cite{gottlieb_etal_2001} , (SSPRK2 and SSPRK3, respectively) are also applied to serve as an reference.
More information and the definition of the absolute error and the relative error can be found in \cite{chu_2018}.

\subsubsection{Sine Wave Streaming}
This test involves the streaming part only: the collision term is turned off. 
A periodic domain $D$ was applied and the initial condition is given by...
More information can be found in \cite{chu_2018}.
In Figure ~\ref{fig: SineWaveStreaming}, the absolute error for the number density is plotted versus the number of elements $N$.
Errors obtained with SSPRK3 and PD-ARS with SSPRK3 decrease as $N^{-3}$ as expected.
All the other schemes have errors decreases as $N^{-2}$.
In the streaming limit, PD-ARSs degenerate to their explicit part, SSPRK2 and SSPRK3, respectively.
Hence, errors of PD-ARS and SSPRKs are indistinguishable on the plot.
Among the second-order schemes, SSP2332 has the smallest error.
\begin{figure}[h]
  \centering
    \includegraphics[width=0.6\textwidth]{figures/SineWaveStreaming}
   \caption{Absolute error versus number of elements $N$ for the streaming sine wave test.  Results employing various time stepping schemes are compared: SSPRK2 (cyan triangles pointing up), SSPRK3 (cyan triangles pointing down), SSP2332 (green crossing), PD-ARS with SSPRK2 (light red circles) and PD-ARS with SSPRK2 (light red hexagram). Black dash-dot reference lines are proportional to $N^{-1}$ (top), $N^{-2}$ (middle), and $N^{-3}$ (bottom), respectively.}
   \label{fig: SineWaveStreaming}
\end{figure}

\subsubsection{Sine Wave Damping}
This test is adapted from \cite{skinnerOstriker_2013}.
Figure~\ref{fig:SineWaveDamping} shows the numerical result of sine wave damping test.
See \cite{chu_2018} for more information about the test setting.
SSP2332 as a second-order scheme display second order convergence rate.
PD-ARSs convergent in first order, as a price paid for realizability-preserving.
\begin{figure}[h]
  \centering
    \includegraphics[width=0.7\textwidth]{figures/SineWaveDamping}
   \caption{Relative error versus number of elements $N$ for the damping sine wave test. Results for different values of the absorption opacity $\sigma_{\Ab}$, employing various IMEX time stepping schemes, are compared.  Errors for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are SSP2332 ($+$), PD-ARS with SSPRK2 (circles) and PD-ARS with SSPRK3 (hexagram).  Black dash-dot reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively.}
  \label{fig:SineWaveDamping}
\end{figure}

\subsubsection{Sine Wave Diffusion}
The last test with known smooth solution is sine wave diffusion test.
It is adopted from \cite{radice_etal_2013}.
See \cite{chu_2018} for more information about the test setting.
Figure~\ref{fig:SineWaveDiffusionJ} shows the numerical result.
SSP2332 and PD-ARSs display third-order accuracy for the number density $\cJ$ and second-oder accuracy for $\cH_{x}$ and their errors are difficult to distinguish.
PD-ARS with SSP2332 behaviors as good as SSP2332 in the diffusion region but requires 1/3 less implicit solver.
\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionJ}
    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionH}
  \end{tabular}
   \caption{Absolute error for the number density $\cJ$ (left) and the number flux $\cH_{x}$ (right) versus number of elements for the sine wave diffusion test.  Results with different values of the scattering opacity $\sigma_{\Scatt}$, employing different IMEX schemes, are compared.  Errors with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are:  SSP2332 ($+$), PD-ARS with SSPRK2 (circles) and PD-ARS with SSPRK3 (hexagram). Black dash-dot on the left plot reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively. Black dash-dot on the right plot reference lines are proportional to $N^{-2}$ .}
   \label{fig:SineWaveDiffusionJ}
\end{figure}

\subsection{Neutrino Stationary State Test} \label{se: Neutrino Stationary State Test}
Next we consider a more realistic-like test: two-dimensional neutrino transport with emission, absorption and elastic scattering under a stationary realistic-like background.
This test is designed to test the realizability-preserving properties of the PD-ARSs schemes.
Figure~\ref{fig:NeutrinoStationaryTestEOS} plots the thermal state and the opacities we used.
This test is computed on a two-dimensional domain $D=\{\vect{x}\in\bbR^{2}:x^{1}\in[0,200], x^{2}\in[0,200]\}$ with a grid of 128 elements on each direction.
We take 10 energy groups vary from 0~MeV to 300~MeV and initial the neutrino number density $\cJ$ and flux density $\bcH$ both zero(arbitrary small).
For realizability-preserving test, CB closure with PD-ARSs schemes, SSP2332, scheme proposed by Cavaglieri \& Bewley \cite{cavaglieriBewley2015} and scheme proposed by McClarren et al. \cite{mcclarren_etal_2008} are all applied.
Only PD-ARSs schemes produces realizability-preserving result and evolve the model to a stable state.
The realizability-preserving plot and the evolve process of PD-ARSs are given in Figure~\ref{fig:NeutrinoStationaryTestEvolve}.

\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.45\textwidth]{figures/NStatinaryS_EOS}
    \includegraphics[width=0.45\textwidth]{figures/NSS_Opacities}
  \end{tabular}
   \caption{Thermal state plot (left) and neutrino absorptivity ($\sigma_{\Ab}$) and neutrino elastic scattering opacity ($ \sigma_{\Scatt}$) in space (right) for the neutrino stationary state test.}
   \label{fig:NeutrinoStationaryTestEOS}
\end{figure}

\begin{figure}[h]
  \centering
    \includegraphics[width=\textwidth]{figures/NSS_1_1}\\
    \includegraphics[width=\textwidth]{figures/NSS_3_1} \\
    \includegraphics[width=\textwidth]{figures/NSS_5_1} \\
   \caption{Readability plots (left column), the number density $\cJ$ versus radius plots (center column) and the flux factor $|\bcH|/\cJ$ versus radius plots (right column) at t = 0.01~ms ,0.35~ms and 5.0~ms for the neutrino stationary state test. For the realizability plots, the light blue area demonstrate the realizability domain and the black lines define its boundary. Each $\bcM=(\cJ,\bcH)^{T}$ state is marked by a red dot. The results of PD-ARS with SSPRK2 and PD-ARS with SSPRK3 are indistinguishable in these plot.}
      \label{fig:NeutrinoStationaryTestEvolve}
\end{figure}

>>>>

In this section we present the numerical results obtained with the PD-ARS schemes given in Section~\ref{se:TimeIntegration}.
The tests in Section \ref{se: Accuracy Tests} are designed to compare the accuracy of the schemes in streaming, absorption, and scattering-dominated regimes in one spatial dimension.
The test in Section \ref{se: Neutrino Stationary State Test} demonstrates the convex-invariance of PD-ARS schemes schemes.
All the tests in this subsection were constructed with third-order accurate spatial discretization (polynomials of degree $k=2$) and time step $\dt = 0.1 \times \dx $.

\subsection{Accuracy Tests}
\label{se: Accuracy Tests}
To compare the accuracy of the IMEX schemes, we applied our PD-ARS schemes and Pareschi \& Russo \cite{pareschiRusso_2005} (SSP2332) scheme to some problems with known smooth solutions in streaming, absorption, ans scattering-dominated regimes in one spatial dimension.
All the tests in this subsection were constructed with the maximum entropy closure in the low-occupancy limit(i.e., the Minerbo closure).
In the streaming test, the second- and third-order accurate explicit strong-stability-preserving Runge-Kutta methods\cite{gottlieb_etal_2001} (SSPRK2 and SSPRK3, respectively) are plotted as references.
To compare the numerical results to analytic solutions, the cell-averaged absolute error or the cell averaged relative error are computed.
For our convergence test, the number of elements ($N$) vary from 8 to 128.

\subsubsection{Sine Wave Streaming}
The sine wave streaming test is designed for the steaming dominated regime, with $\sigma_{\Ab} = \sigma_{\Scatt} = 0$.
A periodic domain with length=1 is used and the initial condition is $\cJ = \bcH =$ a spatial sine wave function.
We evolve the test until the sine wave has completed 10 crossings of the domain.
Figure~\ref{fig: SineWaveStreaming} plots the absolute error for the number density versus the number of elements $N$.
We see the errors obtained with SSPRK3 and PD-ARS with SSPRK3 as the explicit part decrease as $N^{-3}$, as expected.
All the other schemes have errors decreasing as $N^{-2}$.
Among the second-order schemes, SSP2332 has the smallest error.

In the streaming limit, PD-ARS scheme limit to its explicit part, SSPRK2 and SSPRK3, respectively.
That's why the absolute errors of PD-ARS and SSPRK are indistinguishable on the plot.

\begin{figure}[h]
  \centering
    \includegraphics[width=0.8\textwidth]{figures/SineWaveStreaming}
   \caption{Absolute error versus number of elements $N$ for the streaming sine wave test.  Results employing various time stepping schemes are compared: SSPRK2 (cyan triangles pointing up), SSPRK3 (cyan triangles pointing down), SSP2332 (green crosses), PD-ARS with SSPRK2 (light red circles) and PD-ARS with SSPRK2 (light red hexagrams). Black dashed reference lines are proportional to $N^{-2}$ (top), and $N^{-3}$ (bottom), respectively.}
   \label{fig: SineWaveStreaming}
\end{figure}

\subsubsection{Sine Wave Damping}
This test, adapted from \cite{skinnerOstriker_2013}, is designed for absorption-dominated regimes, with $\sigma_{\Scatt} = 0$ and $f_0 = 0$, which results in exponential damping of the wave amplitude.
A periodic domain with length=1 and an initial condition with $\cJ = \bcH =$ a spacial sine wave function is used.
The amplitude of the analytical solution decreases as $e^{-\sigma_{\Ab} t}$.
For $\sigma_{\Ab}$ = 0.1, 1 and 10 we evolve the test until the initial condition has been damped by a factor $e^{-10}$. 
Figure~\ref{fig:SineWaveDamping} shows convergence results of the sine wave damping test in the relative error.
SSP2332, a second-order scheme, displays second-order convergence.
PD-ARS schemes converge with a first-order convergence rate given the price paid for convex-invariance.
\begin{figure}[h]
  \centering
    \includegraphics[width=0.9\textwidth]{figures/SineWaveDamping}
   \caption{Relative error versus number of elements, $N$, for the damping sine wave test. Results for different values of the absorption opacity $\sigma_{\Ab}$, employing various IMEX time stepping schemes, are compared.  Errors for $\sigma_{\Ab}=0.1$, $1$, and $10$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are SSP2332 ($+$), PD-ARS with SSPRK2 (circles) and PD-ARS with SSPRK3 (hexagram).  Black dashed reference lines are proportional to $N^{-1}$ (top) and $N^{-2}$ (bottom), respectively.}
  \label{fig:SineWaveDamping}
\end{figure}

\subsubsection{Sine Wave Diffusion}
The last test with known smooth solution,adopted from \cite{radice_etal_2013}, is the sine wave diffusion test with $\sigma_{\Ab} = 0, f_0 = 0$.
A periodic domain with $D=\{x:x\in[-3,3]\}$ and an initial condition with $\cJ_{0}$ = a spatial sine wave function and $\cH =-\f{1}{3\sigma_{\Scatt}}\pderiv{\cJ}{x}$ is used.
The analytic solution is given by $\cJ = \cJ_{0} \times e^{-\f{\pi^2 t}{27\sigma_{\Scatt}}}$ and $\cH = (3\,\sigma_{\Scatt})^{-1}\pd{\cJ}{x}$.
We evolve with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$, and adjust the end time so that $t_{\mbox{\tiny end}}/\sigma_{\Scatt}=1$. 
Then the amplitude of the sine wave has been reduced by a factor $e^{-\pi^{2}/27}\approx0.694$ for all values of $\sigma_{\Scatt}$. 
Figure~\ref{fig:SineWaveDiffusionJ} shows the convergence.
SSP2332 and PD-ARS schemes display third-order accuracy for the number density, $\cJ$, and second-order accuracy for $\cH_{x}$, and their errors are difficult to distinguish.
PD-ARS with SSP2332 behaves as good as SSP2332 in the diffusion region but requires 1/3 less implicit solvers.
\begin{figure}[h]
  \centering
  \centerline{\includegraphics[width=1.2\textwidth]{figures/SineWaveDiffusion}}
%  \begin{tabular}{cc}
%    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionJ}
%    \includegraphics[width=0.5\textwidth]{figures/SineWaveDiffusionH}
%  \end{tabular}
   \caption{Absolute error for the number density $\cJ$ (left) and the number flux $\cH_{x}$ (right) versus number of elements for the sine wave diffusion test.  Results with different values of the scattering opacity $\sigma_{\Scatt}$, employing different IMEX schemes, are compared.  Errors with $\sigma_{\Scatt}=10^{2}$, $10^{3}$, and $10^{4}$ are plotted with red, green, and blue lines, respectively.  The IMEX schemes employed are:  SSP2332 ($+$), PD-ARS with SSPRK2 (circles), and PD-ARS with SSPRK3 (hexagram). Black dashed lines on the left plot reference lines proportional to $N^{-2}$ (top) and $N^{-3}$ (bottom), respectively. Black dashed line on the right plot reference lines proportional to $N^{-2}$ .}
   \label{fig:SineWaveDiffusionJ}
\end{figure}

\subsection{Neutrino Stationary State Test} \label{se: Neutrino Stationary State Test}
Next we consider a more ``realistic'' test: two-dimensional neutrino transport with emission, absorption, and elastic scattering in a stationary ``realistic'' background.
This is designed to test the realizability-preserving properties of the PD-ARS schemes.
Figure~\ref{fig:NeutrinoStationaryTestEOS} plots the thermal state and the opacities we used.
This test is constructed on a two-dimensional domain $D=\{\vect{x}\in\bbR^{2}:x^{1}\in[0,200], x^{2}\in[0,200]\}$ with a grid of 128 elements in each direction.
We take 10 energy groups from 0~MeV to 300~MeV and initialize the neutrino number density $\cJ$ and flux density $\bcH$ both at zero(arbitrarily small).
For the realizability-preserving test, CB closure with PD-ARS schemes, SSP2332, a scheme proposed by Cavaglieri \& Bewley \cite{cavaglieriBewley2015}, and a scheme proposed by McClarren et al \cite{mcclarren_etal_2008} are all applied.
Only PD-ARS schemes produce realizability-preserving results and evolve the model to a stable state.
The realizability-preserving plot and the evolve process of PD-ARS schemes are given in Figure~\ref{fig:NeutrinoStationaryTestEvolve}.

\begin{figure}[h]
  \centering
  \begin{tabular}{cc}
    \includegraphics[width=0.45\textwidth]{figures/NStatinaryS_EOS}
    \includegraphics[width=0.45\textwidth]{figures/NSS_Opacities}
  \end{tabular}
   \caption{Thermal state (left) and neutrino absorptivity ($\sigma_{\Ab}$) and neutrino elastic scattering opacity ($ \sigma_{\Scatt}$) (right) for the neutrino stationary state test.}
   \label{fig:NeutrinoStationaryTestEOS}
\end{figure}

\begin{figure}[h]
  \centering
    \includegraphics[width=\textwidth]{figures/NSS_1_1}\\
    \includegraphics[width=\textwidth]{figures/NSS_3_1} \\
    \includegraphics[width=\textwidth]{figures/NSS_5_1} \\
    \caption{Realizability (left column), the number density $\cJ$ versus radius (center column), and the flux factor $|\bcH|/\cJ$ versus radius (right column) at t = 0.01~ms ,0.35~ms and 5.0~ms for the neutrino stationary state test. For the realizability plots, the light blue area highlights the realizability domain, and the black lines define its boundary. Each $\bcM=(\cJ,\bcH)^{T}$ state is marked by a red dot. The results of PD-ARS with SSPRK2 and PD-ARS with SSPRK3 are indistinguishable in these plots.}
    \label{fig:NeutrinoStationaryTestEvolve}
\end{figure}

\clearpage

### 2018-10-17 Documents/M1/Astronum_2018/sections/SpacialDiscretization.tex
Modified the IMEX section.
Updated the Astronum_2018 document.

\section{Spacial Discretization}\label{se:SpacialDiscretization}
Here we briefly outline our spacial discretization with the first order discontinuous Galerkin method.
Let the spatial dimension be $d$ and employ Cartesian coordinates.
We divide the spatial domain $D$ evenly into $N$ cells and denote the cell as $\bK_{i}$ with $i = 1,\ldots,N$, so that
\begin{equation*}
D = \cup_{i = 1}^{N} \bK_{i} \quad \text{with} \quad
  \bK_{i}=\{\,\vect{x} : x^{i} \in K^{i} := (\xL^{i},\xH^{i}),~|~i=1,\ldots,d\,\}.
\end{equation*}
Each of the cell has a volume of $|\bK|$
\begin{equation*}
  |\bK| = \int_{\bK}d\vect{x}, \quad\text{where}\quad d\vect{x} = \prod_{i=1}^{d}dx^{i}.  
\end{equation*}
Then we have cell average moment:
\begin{equation}
\bcM_{i} = \dfrac{1}{|\bK|} \int_{\bK_i}\bcM dx.
\end{equation}
Integrate Eq.~\eqref{eq:momentEquations} for each cell $\bK_{i}$ to have
\begin{equation}
\dfrac{d \bcM_{i}}{d t} = - \dfrac{1}{|\bK|} \left( \widehat{\bcF}(\bcM_{i},\bcM_{i+1}) -  \widehat{\bcF}(\bcM_{i-1},\bcM_{i})\right) + \f{1}{\tau}\,\cC(\bcM_{i}),
\end{equation}
where $\widehat{\bcF}(\vect{\cM}_{a},\vect{\cM}_{b})$ is numerical flux and $\f{1}{\tau}\,\cC(\bcM_{i})$ is the integrated collision term.
In this paper we use the simple Lax-Friedrichs flux that given by
\begin{equation}
  \widehat{\bcF}_{LF}(\vect{\cM}_{a},\vect{\cM}_{b})
  =\f{1}{2}\,\big(\,\bcF(\vect{\cM}_{a})+\bcF(\vect{\cM}_{b})-(\,\vect{\cM}_{b}-\vect{\cM}_{a}\,)\,\big).
  \label{eq:Lax-Friedrichs flux}
\end{equation}
Therefore, by treating the flux term explicitly and the collision term implicitly, we have
\begin{align}
\bcM_{i}^{n+1} = \widetilde{\bcM}^{n}_{i} + \f{\dt}{\tau}\,\cC(\bcM^{n+1}_{i})
\label{eq:MomentIMEX}
\end{align}
with
\begin{align}
\widetilde{\bcM}^{n}_{i} 
& = \bcM_{i}^{n} - \dfrac{\dt}{|\bK|} \left( \widehat{\bcF}(\bcM^{n}_{i},\bcM^{n}_{i+1}) -  \widehat{\bcF}(\bcM^{n}_{i-1},\bcM^{n}_{i})\right)\nonumber \\
& = (1-\beta)\bcM_{i}^{n} + \beta\left[ \f{1}{2}\left( \bcM^{n}_{i+1}-\bcF(\bcM^{n}_{i+1})\right)  + \f{1}{2}\left( \bcM^{n}_{i-1}+\bcF(\bcM^{n}_{i-1})\right)\right]
\label{eq:widetildeM}
\end{align}
where $\beta := \frac{\dt}{|\bK|}$.

Assume $\bcM_{i}^{n}$ is realizable for all $i$. 
For Eq.~\eqref{eq:MomentIMEX}, lemma 3 in \cite{chu_etal_2018} says $\bcM^{n+1}_{i}$ remains realizable given $\f{\dt}{\tau} > 0$ and $\widetilde{\bcM}^{n}_{i}$ is realizable.
Substitute the Lax-Friedrichs flux Eq.~\eqref{eq:Lax-Friedrichs flux} in and $\widetilde{\bcM}^{n}_{i}$ becomes Eq.~\eqref{eq:widetildeM}, a convex combination of $\bcM_{i}^{n}$ and the square bracket term if $\beta \in [0,1]$, and it is realizable if $\beta \in [0,1]$ and the square bracket term is realizable.
Given by Lemma 2 in \cite{chu_etal_2018}, the square bracket term is indeed realizable.
Therefore, $\bcM^{n+1}_{i}$ remains realizable given $\beta \in [0,1]$.
Note that $\beta \in [0,1]$ is also the CFL condition of the Forward-Euler method employing on a transport only Boltzmann equation.

>>>>

\section{Spatial Discretization}\label{se:SpacialDiscretization}
Here we briefly outline our spatial discretization using the first-order discontinuous Galerkin method.
Let the spatial dimension be $d$ and employ Cartesian coordinates.
We divide the spatial domain $D$ evenly into $N$ cells and denote the $i$th cell as $\bK_{i}$ with $i = 1,\ldots,N$, so that
\begin{equation*}
D = \cup_{i = 1}^{N} \bK_{i} \quad \text{with} \quad
  \bK_{i}=\{\,\vect{x} : x^{i} \in K^{i} := (\xL^{i},\xH^{i}),~|~i=1,\ldots,d\,\}.
\end{equation*}
Each of the cells has a volume, $|\bK|$, with
\begin{equation*}
  |\bK| = \int_{\bK}d\vect{x}, \quad\text{where}\quad d\vect{x} = \prod_{i=1}^{d}dx^{i}.  
\end{equation*}
Then, we have cell-average moments:
\begin{equation}
\bcM_{i} = \dfrac{1}{|\bK|} \int_{\bK_i}\bcM dx.
\end{equation}
Integrating Eq.~\eqref{eq:momentEquations} for each cell $\bK_{i}$ given
\begin{equation}
\dfrac{d \bcM_{i}}{d t} = - \dfrac{1}{|\bK|} \left( \widehat{\bcF}(\bcM_{i},\bcM_{i+1}) -  \widehat{\bcF}(\bcM_{i-1},\bcM_{i})\right) + \f{1}{\tau}\,\cC(\bcM_{i}),
\label{eq:SemiDiscretizatedMomentEquation}
\end{equation}
where $\widehat{\bcF}(\vect{\cM}_{a},\vect{\cM}_{b})$ is the numerical flux and $\f{1}{\tau}\,\cC(\bcM_{i})$ is the integrated collision term.
In this paper we use the simple Lax-Friedrichs flux, given by
\begin{equation}
  \widehat{\bcF}_{LF}(\vect{\cM}_{a},\vect{\cM}_{b})
  =\f{1}{2}\,\big(\,\bcF(\vect{\cM}_{a})+\bcF(\vect{\cM}_{b})-(\,\vect{\cM}_{b}-\vect{\cM}_{a}\,)\,\big).
  \label{eq:Lax-Friedrichs flux}
\end{equation}
Therefore, by treating the flux term explicitly and the collision term implicitly, we have
\begin{align}
\bcM_{i}^{n+1} = \widetilde{\bcM}^{n}_{i} + \f{\dt}{\tau}\,\cC(\bcM^{n+1}_{i})
\label{eq:MomentIMEX}
\end{align}
with
\begin{align}
\widetilde{\bcM}^{n}_{i} 
& = \bcM_{i}^{n} - \dfrac{\dt}{|\bK|} \left( \widehat{\bcF}(\bcM^{n}_{i},\bcM^{n}_{i+1}) -  \widehat{\bcF}(\bcM^{n}_{i-1},\bcM^{n}_{i})\right)\nonumber \\
& = (1-\beta)\bcM_{i}^{n} + \beta\left[ \f{1}{2}\left( \bcM^{n}_{i+1}-\bcF(\bcM^{n}_{i+1})\right)  + \f{1}{2}\left( \bcM^{n}_{i-1}+\bcF(\bcM^{n}_{i-1})\right)\right]
\label{eq:widetildeM}
\end{align}
where $\beta := \frac{\dt}{|\bK|}$.

Assume $\bcM_{i}^{n}$ is realizable for all $i$. 
For Eq.~\eqref{eq:MomentIMEX}, lemma 3 in \cite{chu_etal_2018} says that $\bcM^{n+1}_{i}$ remains realizable given $\f{\dt}{\tau} > 0$ and given that $\widetilde{\bcM}^{n}_{i}$ stay realizable.
Substituting the Lax-Friedrichs flux, Eq.~\eqref{eq:Lax-Friedrichs flux}, in Eq.~\eqref{eq:SemiDiscretizatedMomentEquation}, $\widetilde{\bcM}^{n}_{i}$ becomes a convex combination of $\bcM_{i}^{n}$ and the square bracket in Eq.~\eqref{eq:widetildeM} if $\beta \in [0,1]$, and it is realizable if $\beta \in [0,1]$ and the term in square brackets is realizable.
Given Lemma 2 in \cite{chu_etal_2018}, the term in square brackets is indeed realizable.
Therefore, $\bcM^{n+1}_{i}$ remains realizable given $\beta \in [0,1]$.
Note that $\beta \in [0,1]$ is also the CFL condition of the Forward-Euler method applied to a transport-only Boltzmann equation.
